**Question 1** 

Normal distribution with a known mean $\mu$, but unknown variance $\sigma^2$

Likelihood: $f(x_1,...x_2|\mu,\sigma^2) = f(x_n|\mu,\sigma^2)=\prod \frac{1}{\sqrt{2\pi\sigma^2}}exp\{\frac{-(x_1-\mu)}{2\sigma^2}\}$ (standard for of normal distribution)

$$f(x_n|\mu, \sigma^2)=(2\pi)^-\frac{2}{n}(\sigma^2)e^-\frac{1}{\sigma^2}\sum\frac{(x_i-\mu)^2}{2}$$ - liklihood with $\sigma^2$ as parameter of interest

$$ Inverse\ Gamma: f(Y|\alpha, beta) = \frac{\beta^\alpha}{\gamma(\alpha)}\cdot y^-(\alpha=1)\cdot e^{(-\beta/y)}$$


whe we think about prior, we write it in terms of sigma 2, so instead of y, we will have sigma2 

$$posterior \propto likelihood * prior $$

ignore constants $$[prior \cdot \frac{\beta^\alpha}{\gamma(\alpha)};/likelihood\ (2\pi)^\frac{-n}{2}]$$

now multiply the prior and likelihood, ignoring the prior:

$$ prior \cdot likelihood \propto (\sigma^2)^{-(\alpha+1)} e^{(\frac{\beta}{\sigma^2})}\cdot (\sigma^2)^{-(\frac{n}{2})}\cdot e^\frac{1}{\sigma^2}\sum\frac{(\mu_i-mu)^2}{2}  $$
group like terms together 










**Question 2** 

**2a): Find the MLE of p (r)**

Process:

To find the MLE, maximize $L(\theta|x)$, with respect to $\theta$:

1. calculate the derivative of $L(\theta|x)$ with respect to $\theta$
2. set the derivative equal to zero
3. Solve the resulting equation for $\theta\$

$$ L(x|p) =\prod_{i=1}{}  {n \choose x}{f(x_i|p)={n \choose x}p^x}(1-p)^{1-x}  $$
Becasue the liklihood function relates only to parameter p, 
 $\frac{n!}{x!(n-x)!}$ is considered to be a fixed constant and is removed from consideration.
 
taking the log-liklihood:

$$ LogL(p|x) = l(p|x) = \sum\limits_{i=1}^n  logp+(n-x)\ log(1-p)$$

$$l`(p|x) = \frac{x}{p}+\frac{n-x}{1-p}(-1) $$

we're looking for the maximizer so we need to set the equation to zero and solve for p
$$l`(p|x) = \frac{x}{p}+\frac{n-x}{1-p}(-1) = 0 $$

$$\frac{x}{p}=\frac{n-x}{1-p} \Rightarrow $$

$$ \hat{p}=\frac{x}{n} $$
$$\hat{p}=\frac{3}{12} = .25$$

**2b) Find the MLE using R's 'optim' function**

https://kevintshoemaker.github.io/NRES-746/LAB3.html
https://stackoverflow.com/questions/37552469/r-estimating-parameters-of-binomial-distribution
https://stackoverflow.com/questions/26734744/how-to-plot-the-log-likelihood-of-binomial-distribution

```{r}
library(MASS)

set.seed(100)

#using Bolkers method--- p = 0.25

k = 3
N = 12

binomNLL1 = function(p, k, N){
  -sum(dbinom(k, prob=p, size = N, log=TRUE))
}

bolkers_approach <- optim(fn=binomNLL1, par= c(p=0.5), N=12, k=k) 
bolkers_approach

```

**2c) Brute force approach**

```{r}
#Negative likelihood function
binomNLL1 <- function(p) -sum(log(dbinom(3, 12, p)))


#create a sequence to test with
p.seq <- seq(0.01, 0.99, 0.01)

#create a test vector 
Likelihood.storage <- 100

#Loop through sequence to find MLE
for(i in p.seq){
  NLL <- (binomNLL1(i))
  if(NLL < Likelihood.storage){
    Likelihood.storage <- NLL
    p <- i
  }
  
}

print(Likelihood.storage)
print(p)

# #or using dplyr
# library(broom)
# library(dplyr)
# library(ggplot2)
# library(readr)
# library(sm)
# library(tidyr)
# 
# 
# crossing(
#   p.seq = seq(0.01, 0.99, 0.01)
# ) %>%
#   rowwise() %>%
#   mutate(
#     NLL = -sum(log(dbinom(3, 12, prob = p.seq))))
#   ) %>%
#   arrange(desc(NLL))
```


**2D - Bayesian Model Using rJAGS**

```{r}
# set.seed(1280)
# 
# #############################
# #
# # Step #1: Set the working directory, and the directory where JAGS lives
# #
# #############################
# 
# jags.directory = "C:/Program Files/JAGS"
# setwd("C:/Users/wmichael/Google Drive/BEE569/labs/test")
# #############################
# #
# # Step #2: Load all the libraries and data
# #
# #############################
# 
# library(R2jags)
# library(abind)
# library(boot)
# 
# #
# # Step #3: Write the statistical model code to a text file
# #
# #############################
# 
# sink("PS2_KoalaModel.jags")
# 
# cat("
# model 
# {
#   x ~ dbin(p, 12) #data sampled binomially with n=12 
#   p ~ dunif(0,1) # prior for the sex ratio of pouch young
# 
# }")
# 
# sink()
# 
# #############################
# #
# # Step #4: Make a list where you include all the data the model will need to run
# #
# #############################
# 
# 
# dat <- list(x = 3)
# 
# #############################
# 
# #############################
# #
# # Step #5: Make a function (with no inputs) where you put a list of parameters and their initial values
# # these are just starting values for your parameters. Hard coded in this case, where m is drawn from a lognormal distribution
# #
# #############################
# 
# InitStage <- function() {list(p = .5)}
# #############################
# #
# # Step #6: Make a column vector with the names of the parameters you want to track
# #
# #############################
# 
# ParsStage <- c("p")
# 
# #############################
# #
# # Step #7: Set the variables for the MCMC
# #
# #############################
# 
# ni <- 10000  # number of draws from the posterior
# nt <- 1    #thinning rate
# nb <- 1000  # number to discard for burn-in
# nc <- 2  # number of chains
# 
# #############################
# #
# # Step #8: Run the jags function to run the code
# #
# #############################
# 
# model = jags(inits=InitStage,
#           n.chains=nc,
#           model.file="PS2_KoalaModel.jags",
#           working.directory=getwd(),
#           data=dat,
#           parameters.to.save=ParsStage,
#           n.thin=nt,
#           n.iter=ni,
#           n.burnin=nb,
#           DIC=T)
# #############################
# #
# # Step #9: Print the summary and explore the object that was returned
# #
# #############################
# 
# model # prints results of the model
# 
# names(model)
# 
# names(model$BUGSoutput)
# 
# #The mean of the posterior distribution
# model.mean <- model$BUGSoutput$mean
# model.sd <- model$BUGSoutput$sd
# 
# model.mean
# model.sd
# 
# #Pull out the credible intervals from the summary output
# model$BUGSoutput$summary[row.names(model$BUGSoutput$summary)=='p',3] #lower
# model$BUGSoutput$summary[row.names(model$BUGSoutput$summary)=='p',7] #upper 

```

