**Question 2** 

**2a): Find the MLE of p (r)**

Process:

To find the MLE, maximize $L(\theta|x)$, with respect to $\theta$:

1. calculate the derivative of $L(\theta|x)$ with respect to $\theta$
2. set the derivative equal to zero
3. Solve the resulting equation for $\theta\$

$$ L(x|p) =\prod_{i=1}{}  {n \choose x}{f(x_i|p)={n \choose x}p^x}(1-p)^{1-x}  $$
Becasue the liklihood function relates only to parameter p, 
 $\frac{n!}{x!(n-x)!}$ is considered to be a fixed constant and is removed from consideration.
 
taking the log-liklihood:

$$ LogL(p|x) = l(p|x) = \sum\limits_{i=1}^n  logp+(n-x)\ log(1-p)$$

$$l`(p|x) = \frac{x}{p}+\frac{n-x}{1-p}(-1) $$

we're looking for the maximizer so we need to set the equation to zero and solve for p
$$l`(p|x) = \frac{x}{p}+\frac{n-x}{1-p}(-1) = 0 $$

$$\frac{x}{p}=\frac{n-x}{1-p} \Rightarrow $$

$$ \hat{p}=\frac{x}{n} $$
$$\hat{p}=\frac{3}{12} = .25$$

**2b) Find the MLE using R's 'optim' function**

https://kevintshoemaker.github.io/NRES-746/LAB3.html
https://stackoverflow.com/questions/37552469/r-estimating-parameters-of-binomial-distribution
https://stackoverflow.com/questions/26734744/how-to-plot-the-log-likelihood-of-binomial-distribution

```{r}
library(MASS)

set.seed(100)

#using Bolkers method--- p = 0.25

k = 3
N = 12

binomNLL1 = function(p, k, N){
  -sum(dbinom(k, prob=p, size = N, log=TRUE))
}

bolkers_approach <- optim(fn=binomNLL1, par= c(p=0.5), N=12, k=k) 
bolkers_approach

```

**2c) Brute force approach**

```{r}
binomNLL1 <- function(p) -sum(log(dbinom(3, 12, p)))

#create a sequence to test with
p.seq <- seq(0.01, 0.99, 0.01)

#visually plot 
plot(p.seq, sapply(p.seq, binomNLL1), type="l")

x <- sapply(p.seq, binomNLL1)
```

