---
title: "ps03_michael_wethington"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Question 1** 

$$Beta(\alpha, \beta) = \frac{\theta^{a-1}(1-\theta)^{b-1}}{\beta(a,b)} = constant \cdot \theta^{a-1}(1-\theta)^{b-1}$$
$$ P(\theta|data) = \frac{P(data|\theta)\cdot P(\theta)}{P(data)} \propto P(data|\theta)\cdot P(\theta) $$

For our likelihood, we insert the binomial distribution  
$$ P(data|\theta) \propto \theta^z(1-\theta)^{N-z}; z = \sum x_i$$
Multiplying the likelihood by our prior, the beta distribution, we find our posterior distribution, in the form of the beta distribution.  

$$P(\theta|data \propto \theta^{a+z-1}(1-\theta)^{N+b-z-1}$$

Where $$a'=a+z$$
$$b'=N+b-z$$
$$P(\theta|data)= \frac{\theta^{a'-1}(1-\theta)^{b'-1}}{\beta(a',b')} $$






**Question 2** 

Because the parameter of the Poisson distribution can in principle be any positive real number, we want a prior whose support is $(0,\infty)$. 

The posterior with the conjugate gamma prior:


Likelihood of poisson 
$$x_i~P_o(\lambda)\Rightarrow P(x_i|\lambda)=\frac{\lambda^{x_i}e^{-\lambda}}{x_i!}$$
we assume lambda is gamma distributed with parameters $\alpha, \beta$

$$\lambda ~Gamma(\alpha,\beta)\Rightarrow P(\lambda) = \frac{\beta^{\alpha}}{\Gamma(\alpha)}\lambda^{\alpha-1}e^{-\beta\lambda} $$ 
$$P(\lambda|x_i,...x_n)= \frac{P(x|\lambda)P(\lambda)}{P (X)}$$
$$\propto P(x|\lambda)P(\lambda)$$
Assuming we have independent observations, we can ignore the normalizing constant, and find the likelihood  

$$P(x|\lambda) = \prod \frac{\lambda^{x_i}e^{-\lambda}}{x_i!} = /frac{\lambda^{x_1 + x_2+...x_n}e^{-N\lambda}}$$
$$\sum x_i = N\bar{x}$$
$$P(\lambda|x) \propto P(x|\lambda)P(\lambda)=\lambda^{N\bar{x}}e^{-N\lambda}\lambda^{\alpha-1}e^{-\beta\lambda}$$
$$= \lambda^{N\bar{x}+\alpha-1}e^{-(\beta+N)\lambda} ~Gamma(N\bar{x}+\alpha,  \beta+N) $$





**Question 3** 


```{r}
jags.directory = "C:/Program Files/JAGS"
setwd("C:/Users/wmichael/Google Drive/BEE569/problem_sets/03/scripts")

library(R2jags)
library(abind)
library(boot)
library(ggplot2)
library(cowplot) #library for multiplots
library(MASS) #for fitdistr function

#################################################
#Uniform Model Builder
#########################
# sink("Pois_Unif1.jags")
# 
# cat("
# model {
#   for (i in 1:10) #Loop through each data entry
#   {
# 		y[i] ~ dpois(lambda) #define likelihood
#   }
# 
# 	#Define Prior
# 	lambda ~ dunif(0, 10)
# 
# }")
# sink()
# 
# 
# sink("Pois_Unif2.jags")
# 
# cat("
# model {
#   for (i in 1:10) #Loop through each data entry
#   {
# 		y[i] ~ dpois(lambda) #define likelihood
#   }
# 
# 	#Define Prior
# 	lambda ~ dunif(0, 20)
# 
# }")
# sink()
# 
# 
# 
# 
# sink("Pois_Unif3.jags")
# cat("
# model {
#   for (i in 1:10) #Loop through each data entry
#   {
# 		y[i] ~ dpois(lambda) #define likelihood
#   }
# 
# 	#Define Prior
# 	lambda ~ dunif(5, 15)
# }")
# sink()
# 
# 
# #################################################
# #Gamma Model Builder
# #########################
# sink("Pois_Gamma1.jags")
# 
# cat("
# model {
#   for (i in 1:10) #Loop through each data entry
#   {
# 		y[i] ~ dpois(lambda) #define likelihood
#   }
# 
# 	#Define Prior
# 	lambda ~ dgamma(37, 100)
# 
# }")
# sink()
# 
# 
# sink("Pois_Gamma2.jags")
# 
# cat("
# model {
#   for (i in 1:10) #Loop through each data entry
#   {
# 		y[i] ~ dpois(lambda) #define likelihood
#   }
# 
# 	#Define Prior
# 	lambda ~ dgamma(6, 2)
# 
# }")
# sink()
# 
# 
# 
# 
# sink("Pois_Gamma3.jags")
# 
# cat("
# model {
#   for (i in 1:10) #Loop through each data entry
#   {
# 		y[i] ~ dpois(lambda) #define likelihood
#   }
# 
# 	#Define Prior
# 	lambda ~ dgamma(30, 2)
# 
# }")
# sink()
# 
# 
# ################################################
# #Lognormal Model Builder
# #########################
# sink("Pois_LgNorm1.jags")
# 
# cat("
# model {
#   for (i in 1:10) #Loop through each data entry
#   {
# 		y[i] ~ dpois(lambda) #define likelihood
#   }
# 
# 	#Define Prior
# 	lambda ~ dlnorm(1, 1)
# 	
# 
# }")
# sink()
# 
# 
# sink("Pois_LgNorm2.jags")
# 
# cat("
# model {
#   for (i in 1:10) #Loop through each data entry
#   {
# 		y[i] ~ dpois(lambda) #define likelihood
#   }
# 
# 	#Define Prior
# 	lambda ~ dlnorm(4, 10)
# 	
# 
# }")
# sink()
# 
# 
# sink("Pois_LgNorm3.jags")
# 
# cat("
# model {
#   for (i in 1:10) #Loop through each data entry
#   {
# 		y[i] ~ dpois(lambda) #define likelihood
#   }
# 
# 	#Define Prior
# 	lambda ~ dlnorm(2, 20)
# 	
# 
# }")
# sink()
# 
# 




#################################################
#Data and parameter specification
#################################################
set.seed(1001)
Dat <- list(y = c(3,5,2,3,3,0,2,6,1,3))
probs = c(0.025, 0.95) # define intervals


unif.inits <- function() {list(lambda = runif(1,0,10))} #trying to initialize 1 lambda for each chain
unif.inits2 <- function() {list(lambda = runif(1,0,20))} #trying to initialize 1 lambda for each chain
unif.inits3 <- function() {list(lambda = runif(1,5,15))} #trying to initialize 1 lambda for each chain
Pois.inits <- function() {list(lambda = runif(1,0,10))} #trying to initialize 1 lambda for each chain
lgNorm.inits <- function() {list(lambda = runif(1,0,10))} #trying to initialize 1 lambda for each chain


Params <- c("lambda")


ni <- 1100  # number of draws from the posterior
nt <- 1    #thinning rate (reduces autocorrelation)
nb <- 100  # number to discard for burn-in (when parameter start wide, and start moving close to each other)
nc <- 3  # number of chains


#################################################
#Uniform Models Compiler
################################################
Pois_Unif1 = jags(inits=unif.inits,
                  n.chains=nc,
                  model.file="Pois_Unif1.jags",
                  working.directory=getwd(),
                  data=Dat,
                  parameters.to.save=Params,
                  n.thin=nt,
                  n.iter=ni,
                  n.burnin=nb,
                  DIC=T)


Pois_Unif2 = jags(inits=unif.inits2,
                  n.chains=nc,
                  model.file="Pois_Unif2.jags",
                  working.directory=getwd(),
                  data=Dat,
                  parameters.to.save=Params,
                  n.thin=nt,
                  n.iter=ni,
                  n.burnin=nb,
                  DIC=T)

Pois_Unif3 = jags(inits=unif.inits3,
                  n.chains=nc,
                  model.file="Pois_Unif3.jags",
                  working.directory=getwd(),
                  data=Dat,
                  parameters.to.save=Params,
                  n.thin=nt,
                  n.iter=ni,
                  n.burnin=nb,
                  DIC=T)

################################################
#Gamma Models Compiler
################################################
Pois_Gamma1 = jags(inits=Pois.inits,
                   n.chains=nc,
                   model.file="Pois_Gamma1.jags",
                   working.directory=getwd(),
                   data=Dat,
                   parameters.to.save=Params,
                   n.thin=nt,
                   n.iter=ni,
                   n.burnin=nb,
                   DIC=T)

Pois_Gamma2 = jags(inits=Pois.inits,
                   n.chains=nc,
                   model.file="Pois_Gamma2.jags",
                   working.directory=getwd(),
                   data=Dat,
                   parameters.to.save=Params,
                   n.thin=nt,
                   n.iter=ni,
                   n.burnin=nb,
                   DIC=T)

Pois_Gamma3 = jags(inits=Pois.inits,
                   n.chains=nc,
                   model.file="Pois_Gamma3.jags",
                   working.directory=getwd(),
                   data=Dat,
                   parameters.to.save=Params,
                   n.thin=nt,
                   n.iter=ni,
                   n.burnin=nb,
                   DIC=T)




################################################
#LogNormal Models Compiler
################################################
Pois_LgNorm.jags1 = jags(inits=lgNorm.inits,
                         n.chains=nc,
                         model.file="Pois_LgNorm1.jags",
                         working.directory=getwd(),
                         data=Dat,
                         parameters.to.save=Params,
                         n.thin=nt,
                         n.iter=ni,
                         n.burnin=nb,
                         DIC=T)


Pois_LgNorm.jags2 = jags(inits=lgNorm.inits,
                         n.chains=nc,
                         model.file="Pois_LgNorm2.jags",
                         working.directory=getwd(),
                         data=Dat,
                         parameters.to.save=Params,
                         n.thin=nt,
                         n.iter=ni,
                         n.burnin=nb,
                         DIC=T)

Pois_LgNorm.jags3 = jags(inits=lgNorm.inits,
                         n.chains=nc,
                         model.file="Pois_LgNorm3.jags",
                         working.directory=getwd(),
                         data=Dat,
                         parameters.to.save=Params,
                         n.thin=nt,
                         n.iter=ni,
                         n.burnin=nb,
                         DIC=T)

################################################
#Graph the Uniform Models
################################################
Pois_Unif1.dat <- as.data.frame(Pois_Unif1$BUGSoutput$sims.matrix) # Extract values and Cast to dataframe so we can access the lambda values
#hist(Pois_Unif1.dat$lambda)

Pois_Unif2.dat <- as.data.frame(Pois_Unif2$BUGSoutput$sims.matrix) # Extract values and Cast to dataframe so we can access the lambda values
#hist(Pois_Unif2.dat$lambda)

Pois_Unif3.dat <- as.data.frame(Pois_Unif3$BUGSoutput$sims.matrix) # Extract values and Cast to dataframe so we can access the lambda values
#hist(Pois_Unif3.dat$lambda)


PoisUnif_Pst.plot <-  ggplot(Pois_Unif1.dat, aes(x=lambda, color="#9999")) +
  
  geom_histogram(fill="#d7191c", color="#e9ecef", alpha = 0.8, bins = 30)+
  ggtitle("Poission-Uniform Posterior (0,10)")+
  theme_minimal()+
  theme(plot.title = element_text(size = 8, face = "bold"))


PoisUnif_Pst.plot2 <-  ggplot(Pois_Unif2.dat, aes(x=lambda, color="#9999")) +
  
  geom_histogram(fill="#fdae61", color="#e9ecef", alpha = 0.8, bins = 30)+
  ggtitle("Poission-Uniform Posterior (0,20)")+
  theme_minimal()+
  theme(plot.title = element_text(size = 8, face = "bold"))


PoisUnif_Pst.plot3 <-  ggplot(Pois_Unif3.dat, aes(x=lambda, color="#9999")) +
  
  geom_histogram(fill="#91bfdb", color="#e9ecef", alpha = 0.8, bins = 30)+
  ggtitle("Poission-Uniform Posterior (5,15)")+
  theme_minimal()+
  theme(plot.title = element_text(size = 8, face = "bold"))


#Plot all three uniform prior models together 
unif.posterior <-  ggplot(Pois_Unif1.dat, aes(x=lambda, color="#9999")) +
  
  geom_histogram(fill="#d7191c", color="#d7191c", alpha = 0.8, bins = 30)+
  geom_histogram(data= Pois_Unif2.dat, fill="#fdae61", color="#e9ecef", alpha = 0.8, bins = 30)+
  geom_histogram(data = Pois_Unif3.dat, fill="#91bfdb", color="#e9ecef", alpha = 0.8, bins = 30)+
  ggtitle("Poission-Uniform Posteriors")+
  theme_minimal()

unif.posterior

#Plot all three uniform prior models in a grid
plot_grid(PoisUnif_Pst.plot, PoisUnif_Pst.plot2, PoisUnif_Pst.plot3, labels= "AUTO")



#################################################
#Graph the Gamma Models
#################################################
Pois_Gamma1.dat <- as.data.frame(Pois_Gamma1$BUGSoutput$sims.matrix) # Extract values and Cast to dataframe so we can access the lambda values
#hist(Pois_Gamma1.dat$lambda)

Pois_Gamma2.dat <- as.data.frame(Pois_Gamma2$BUGSoutput$sims.matrix) # Extract values and Cast to dataframe so we can access the lambda values
#hist(Pois_Gamma2.dat$lambda)

Pois_Gamma3.dat <- as.data.frame(Pois_Gamma3$BUGSoutput$sims.matrix) # Extract values and Cast to dataframe so we can access the lambda values
#hist(Pois_Gamma2.dat$lambda)



PoisGamma_Pst.plot <-  ggplot(Pois_Gamma1.dat, aes(x=lambda, color="#9999")) +
  
  geom_histogram(fill="#d7191c", color="#e9ecef", alpha = 0.8, bins = 30)+
  ggtitle("Poission-Gamma Posterior (37,100)")+
  theme_minimal()+
  theme(plot.title = element_text(size = 8, face = "bold"))


PoisGamma_Pst.plot2 <-  ggplot(Pois_Gamma2.dat, aes(x=lambda, color="#9999")) +
  
  geom_histogram(fill="#fdae61", color="#e9ecef", alpha = 0.8, bins = 30)+
  ggtitle("Poission-Gamma Posterior (6,2)")+
  theme_minimal()+
  theme(plot.title = element_text(size = 8, face = "bold"))


PoisGamma_Pst.plot3 <-  ggplot(Pois_Gamma3.dat, aes(x=lambda, color="#9999")) +
  
  geom_histogram(fill="#91bfdb", color="#e9ecef", alpha = 0.8, bins = 30)+
  ggtitle("Poission-Gamma Posterior (30,2)")+
  theme_minimal()+
  theme(plot.title = element_text(size = 8, face = "bold"))


plot_grid(PoisGamma_Pst.plot, PoisGamma_Pst.plot2, PoisGamma_Pst.plot3, labels= "AUTO")




gamma.posterior <-  ggplot(Pois_Gamma1.dat, aes(x=lambda, color="#9999")) +
  
  geom_histogram(fill="#d7191c", color="#d7191c", alpha = 0.8, bins = 30)+
  geom_histogram(data= Pois_Gamma2.dat, fill="#fdae61", color="#e9ecef", alpha = 0.8, bins = 30)+
  geom_histogram(data = Pois_Gamma3.dat, fill="#91bfdb", color="#e9ecef", alpha = 0.8, bins = 30)+
  ggtitle("Poission-Gamma Posteriors")+
  theme_minimal()+
  theme(plot.title = element_text(size = 8, face = "bold"))


gamma.posterior



#################################################
#Graph the Lognormal models
#################################################
#Compile the posteriors
Pois_LgNorm.jags1.dat <- as.data.frame(Pois_LgNorm.jags1$BUGSoutput$sims.matrix) # Extract values and Cast to dataframe so we can access the lambda values
Pois_LgNorm.jags2.dat <- as.data.frame(Pois_LgNorm.jags2$BUGSoutput$sims.matrix) # Extract values and Cast to dataframe so we can access the lambda values
Pois_LgNorm.jags3.dat <- as.data.frame(Pois_LgNorm.jags3$BUGSoutput$sims.matrix) # Extract values and Cast to dataframe so we can access the lambda values


#Plot all three uniform prior models together 
LgNorm.posterior <-  ggplot(Pois_LgNorm.jags1.dat, aes(x=lambda, color="#9999")) +
  
  geom_histogram(fill="#d7191c", color="#d7191c", alpha = 0.8, bins = 30)+
  geom_histogram(data= Pois_LgNorm.jags2.dat, fill="#fdae61", color="#e9ecef", alpha = 0.8, bins = 30)+
  geom_histogram(data = Pois_LgNorm.jags3.dat, fill="#91bfdb", color="#e9ecef", alpha = 0.8, bins = 30)+
  ggtitle("Poission-Lognormal Posteriors")+
  theme_minimal()

LgNorm.posterior


Pois_LgNorm.jags1.plot <-  ggplot(Pois_LgNorm.jags1.dat, aes(x=lambda, color="#9999")) +
  
  geom_histogram(fill="#d7191c", color="#e9ecef", alpha = 0.8, bins = 30)+
  ggtitle("Poission-Lognormal (Posterior (1,1)")+
  theme_minimal() + 
  theme(plot.title = element_text(size = 8, face = "bold"))

Pois_LgNorm.jags2.plot2 <-  ggplot(Pois_LgNorm.jags2.dat, aes(x=lambda, color="#9999")) +
  
  geom_histogram(fill="#fdae61", color="#e9ecef", alpha = 0.8, bins = 30)+
  ggtitle("Poission-Lognormal(Posterior (4,10)")+
  theme_minimal()+
  theme(plot.title = element_text(size = 8, face = "bold"))


Pois_LgNorm.jags3.plot3 <-  ggplot(Pois_LgNorm.jags3.dat, aes(x=lambda, color="#9999")) +
  
  geom_histogram(fill="#91bfdb", color="#e9ecef", alpha = 0.8, bins = 30)+
  ggtitle("Poission-Lognormal Posterior (2,20)")+
  theme_minimal()+
  theme(plot.title = element_text(size = 8, face = "bold"))




#################################################
#Master Plot
#############################
plot_grid(PoisUnif_Pst.plot, PoisUnif_Pst.plot2, PoisUnif_Pst.plot3,
          PoisGamma_Pst.plot, PoisGamma_Pst.plot2, PoisGamma_Pst.plot3,
          Pois_LgNorm.jags1.plot, Pois_LgNorm.jags2.plot2, Pois_LgNorm.jags3.plot3,
          labels= "AUTO", label_size = 6)


########
#Table of Confidence Intervals
#################################################

u1<- quantile(Pois_Unif1.dat$lambda, probs = probs)
u2 <-quantile(Pois_Unif2.dat$lambda, probs = probs)
u3 <- quantile(Pois_Unif3.dat$lambda, probs = probs)


g1 <- quantile(Pois_Gamma1.dat$lambda, probs = probs)
g2 <- quantile(Pois_Gamma2.dat$lambda, probs = probs)
g3 <- quantile(Pois_Gamma3.dat$lambda, probs = probs)


l1 <- quantile(Pois_LgNorm.jags1.dat$lambda, probs = probs)
l2 <- quantile(Pois_LgNorm.jags2.dat$lambda, probs = probs)
l3 <- quantile(Pois_LgNorm.jags3.dat$lambda, probs = probs)



Name.list <- c("Unif1", "Unif2", "Unif3", "Gamma1", "Gamma2","Gamma3", "LogNorm1", "LogNorm2", "LogNorm3")
ci.2.5 <- c(u1[1], u2[1], u3[1], g1[1], g2[1], g2[1], l1[1], l2[1], l3[1])
ci.95 <- c(u1[2], u2[2], u3[2], g1[2], g2[2], g2[2], l1[2], l2[2], l3[2])

df <- data.frame(Name.list, ci.2.5, ci.95 )
 

knitr::kable(df, col.names = gsub("[.]", " ", names(df)))

#Calculate the MLE of the poisson data and approximate the 95% confidence intervals

data <- c(3,5,2,3,3,0,2,6,1,3)
mle <- fitdistr(data, densfun = "Poisson")
se <- as.numeric(mle[2])
MLE.95.CI <- as.numeric(var(data))+(1.96*se)
MLE.2.5.CI <- as.numeric(var(data))-(1.96*se)

#Upper CI
MLE.95.CI

#Lower CI
MLE.2.5.CI

```




