---
title: "Lab06"
author: "Michael Wethington"
date: "9/30/2020"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)


# Step #1: Set the working directory, and the directory where JAGS lives
jags.directory = "C:/Program Files/JAGS"
setwd("C:/Users/wmichael/Google Drive/BEE569/problem_sets/06/scripts")

#write the data
sim.data <- rbeta(100, 2,5)
#write.csv(sim.data, file = "C:/Users/wmichael/Google Drive/BEE569/problem_sets/06/lab06.csv")



# Step #2: Load all the libraries and data
library(R2jags)
library(abind)
library(boot)
library(MASS)

sim.data.100<-rbeta(100,2,5)
sim.data.50<-rbeta(50,2,5)
sim.data.10 <-rbeta(10,2,5)


```

**Exercise 1** 
```{r}
#visualize the data
boxplot(df$Percent ~ df$Group)
```


```{r}



sink("Week6_BetaModel.jags")
cat("
model {
  for (i in 1:length(y)) {
		y[i] ~ dbeta(alpha, beta)
  }
	
	alpha ~ dgamma(.01, .01) #simulate from these values
	beta ~ dgamma(.01, .01)
	
#use the arrow when you want to explicitly define a value (i.e. this value is 3)
	
}",fill = TRUE)
sink()



#create a data list that gets fed to jags
Dat <- list(y = sim.data.100)

#Set initial parameters
InitStage <- function() {list(alpha = 1, beta = 1)}

#Vector for variables to track
ParsStage <- c("alpha", "beta")

#Set up the variable values for the MCMC
ni <- 1100  # number of draws from the posterior
nt <- 1    #thinning rate
nb <- 100  # number to discard for burn-in
nc <- 2  # number of chains

# Step #8: Run the jags function to run the code
m = jags(inits=InitStage,
         n.chains=nc,
         model.file="Week6_BetaModel.jags",
         working.directory=getwd(),
         data=Dat,
         parameters.to.save=ParsStage,
         n.thin=nt,
         n.iter=ni,
         n.burnin=nb,
         DIC=T)

#print out the results
m

#plot the posteriors for alpha and beta
par(mfrow = c(2,1))
hist(m$BUGSoutput$sims.matrix[,1], breaks = 20, main = NULL, xlab = expression(paste("Posterior Distribution of ", alpha, sep = "")))
hist(m$BUGSoutput$sims.matrix[,2], breaks = 20, main = NULL, xlab = expression(paste("Posterior Distribution of ", beta, sep = "")))

#Extract the means (these should be close the the parameter values we initially set (i.e. 2,5)) 
mean(m$BUGSoutput$sims.matrix[,1])
mean(m$BUGSoutput$sims.matrix[,2])



```

**Exercise 1b - Check with different sample totals (n = 50, n = 10). What Happens to the posterior?**






**Exercise 2 - Confirm your results with fitdistr**
```{r}
modl.fit <- fitdistr(sim.data.100, densfun = "beta", start = list(shape1 = 1, shape2 = 1))
mod.fit
```

#Exercise 3

```{r}

  
```

```{r}

```


**Exercise 3 Model the sea slug data with a normal, then with a beta distribution** 



Set working directory, load libraries and raw data
```{r}

jags.directory = "C:/Program Files/JAGS"
setwd("C:/Users/wmichael/Google Drive/BEE569/problem_sets/06/scripts")
df <- read.csv("C:/Users/wmichael/Google Drive/BEE569/problem_sets/06/scripts/seaslugs.csv", header = TRUE)

library(R2jags)
library(abind)
library(boot)
```
Model as a normal $$Y_i \backsim ~ N(\mu_i, \sigma^2)$$

$$y_i |g_i, \mu, \sigma^2 \backsim N (\mu_{gi}, \sigma^2) \\ 
g_i element (1,...G)$$
```{r}

sink("Week6_ex3a.jags")
cat("
model {

  #priors
  sigma ~ dunif(0,100)
  tau <-  pow(sigma, -2)
  

  for (i in 1:length(y)) {
  	y[i] ~ dnorm(means[groups[i]], tau)
  }
  
  
  for (i in 1:8)	{
    means[i] ~ dnorm(0.1, 0.001)
  }

}", fill = TRUE)
sink()

#create a data list that gets fed to jags
Dat <- list(
  y = df$Percent,
  groups = df$Group
)

#create a function for the initial parameters
InitStage <- function() {list(means= rnorm(8,0,0),
                              sigma = rgamma(1,1.0,1.0))}

#create a vector for parameters we want to track
ParsStage <- c("means","sigma")

#Establish MCMC variable 
ni <- 11000  # number of draws from the posterior
nt <- 1    #thinning rate
nb <- 1000  # number to discard for burn-in
nc <- 3  # number of chains

#Compile and run the model
m = jags(inits=InitStage,
         n.chains=nc,
         model.file="Week6_ex3a.jags",
         working.directory=getwd(),
         data=Dat,
         parameters.to.save=ParsStage,
         n.thin=nt,
         n.iter=ni,
         n.burnin=nb,
         DIC=T)

#print the results
m

#pull out the individual chains of data (should have 9 columns, 8 means and 1 sigma)
chains <- m$BUGSoutput$sims.matrix[,2:10]
chains


# Check convergence for group means
par(mfrow = c(4,2))
for(i in 1:8){
  plot(chains[,i],type= "l", main = "Chain Values")
}


#Check convergence for sigma
plot(chains[,9], type = 'l', main = expression(paste("Chain values for Sigma")))

#plot the posterior density distribution for the chains
plot.new()
par(mfrow = c(4,2), mar = rep(2,4))
for (i in 1:8){
  hist(chains[,i], breaks = 30, main = "Posterior Probability Density - Chains", sep = "", col = "lightblue")
}

#plot the posterior density distribution for sigma
plot.new()
par(mfrow = c(4,2), mar = rep(2,4))
hist(chains[,9], breaks = 30, main = "Posterior Probability Density - Sigma", sep = "", col = "lightblue")



```

**Exercise 3b - do it again but with a more appropriate distribution**

Model the data using a Beta distribution because we're working with data that ranges from 0 to 1. 

In order to solve for the mean and variances, we need to use moment method estimation for the Beta distribution



$$Y_i \backsim ~ Beta(\alpha, \beta)$$



$$a = (\frac{1-\mu}{\sigma^2}-\frac{1}{\mu})$$

$$B = \alpha(\frac{1}{\mu}-1)$$


```{r}
sink("Week6_ex3b.jags")
cat("
model {

  #priors
  sigma ~ dunif(0,100)
  tau <-  pow(sigma, -2)
  

  for (i in 1:length(y)) {
  	y[i] ~ dbeta(alpha[groups[i]], alpha[groups[i]])
  }
  
  #loop through and establish group priors
  for (i in 1:8)	{
  alpha[i] ~ dgamma(1,0.001)
  beta[i] ~ dgamma(1,0.001)
  
  #calculate mu and sigma with moment matching
  mu[i] <- alpha[i] / (alpha[i]+beta[i])
  sigma[i] <-pow(alpha[i]*beta[i] / (pow(alpha[i]+beta[i],2))*(alpha[i] + beta[i] + 1),0.5)
  }

}", fill = TRUE)
sink()

#create a data list that gets fed to jags

Dat <- list(
  y = df$Percent,
  groups = df$Group
)

#create a function for the initial parameters
InitStage <- function() {list(alpha = rgamma(8, 0.01, 0.01)), beta = rgamma(8, 0.01, 0.01))}

#create a vector for parameters we want to track
ParsStage <- c("mu","sigma")

#Establish MCMC variable 
ni <- 11000  # number of draws from the posterior
nt <- 1    #thinning rate
nb <- 1000  # number to discard for burn-in
nc <- 3  # number of chains

#Compile and run the model
m = jags(inits=InitStage,
         n.chains=nc,
         model.file="Week6_ex3b.jags",
         working.directory=getwd(),
         data=Dat,
         parameters.to.save=ParsStage,
         n.thin=nt,
         n.iter=ni,
         n.burnin=nb,
         DIC=T)

#print the results
m

#pull out the individual chains of data (should have 9 columns, 8 means and 1 sigma)
chains <- m$BUGSoutput$sims.matrix[,2:10]
chains


# Check convergence for group means
par(mfrow = c(4,2))
for(i in 1:8){
  plot(chains[,i],type= "l", main = "Chain Values")
}


#Check convergence for sigma
plot(chains[,9], type = 'l', main = expression(paste("Chain values for Sigma")))

#plot the posterior density distribution for the chains
plot.new()
par(mfrow = c(4,2), mar = rep(2,4))
for (i in 1:8){
  hist(chains[,i], breaks = 30, main = "Posterior Probability Density - Chains", sep = "", col = "lightblue")
}

#plot the posterior density distribution for sigma
plot.new()
par(mfrow = c(4,2), mar = rep(2,4))
hist(chains[,9], breaks = 30, main = "Posterior Probability Density - Sigma", sep = "", col = "lightblue")



```

**Exercise 4 - Model using contrasts**

Mean of group 1 is the control. 



```{r}

sink("Week6_ex4.jags")
cat("
model {
  #priors
  sigma ~ dunif(0,100)
  tau <-  pow(sigma, -2)
  

  for (i in 1:length(y)) {
  	y[i] ~ dbeta(alpha[groups[i]], alpha[groups[i]])
  }
  
  #loop through and establish group priors
  for (i in 1:8)	{
  alpha[i] ~ dgamma(1,0.001)
  beta[i] ~ dgamma(1,0.001)
  
  #calculate mu and sigma with moment matching
  mu[i] <- alpha[i] / (alpha[i]+beta[i])
  sigma[i] <-pow(alpha[i]*beta[i] / (pow(alpha[i]+beta[i],2))*(alpha[i] + beta[i] + 1),0.5)
  }
  
  #Esbilsh the contrasts to compare against the control group
  for (i in 1:7){
  contrast[i] <- mu[i+1] - mu[i]
  }
}", fill = TRUE)
sink()

#create a data list that gets fed to jags

Dat <- list(
  y = df$Percent,
  groups = df$Group
)

#create a function for the initial parameters
InitStage <- function() {list(alpha = rgamma(8, 0.01, 0.01)), beta = rgamma(8, 0.01, 0.01))}

#create a vector for parameters we want to track
ParsStage <- c("mu","sigma","contrast")

#Establish MCMC variable 
ni <- 11000  # number of draws from the posterior
nt <- 1    #thinning rate
nb <- 1000  # number to discard for burn-in
nc <- 3  # number of chains

#Compile and run the model
m = jags(inits=InitStage,
         n.chains=nc,
         model.file="Week6_ex4.jags",
         working.directory=getwd(),
         data=Dat,
         parameters.to.save=ParsStage,
         n.thin=nt,
         n.iter=ni,
         n.burnin=nb,
         DIC=T)

#print the results
m

#pull out the individual chains of data (should have 9 columns, 8 means and 1 sigma)
chains <- m$BUGSoutput$sims.matrix[,2:10]
chains


# Check convergence for group means
par(mfrow = c(4,2))
for(i in 1:8){
  plot(chains[,i],type= "l", main = "Chain Values")
}


#Check convergence for sigma
plot(chains[,9], type = 'l', main = expression(paste("Chain values for Sigma")))

#plot the posterior density distribution for the chains
plot.new()
par(mfrow = c(4,2), mar = rep(2,4))
for (i in 1:8){
  hist(chains[,i], breaks = 30, main = "Posterior Probability Density - Chains", sep = "", col = "lightblue")
}

#plot the posterior density distribution for sigma
plot.new()
par(mfrow = c(4,2), mar = rep(2,4))
hist(chains[,9], breaks = 30, main = "Posterior Probability Density - Sigma", sep = "", col = "lightblue")





```


