---
title: "PS_07"
author: "Michael Wethington"
date: "10/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r }

library(MCMCvis)
library(R2jags)
library(abind)
library(boot)
library(tidyr)
library(ggplot2)

jags.directory = "C:/Program Files/JAGS"
setwd("C:/Users/wmichael/Google Drive/BEE569/problem_sets/07/data")
data <- read.csv(file = "C:/Users/wmichael/Google Drive/BEE569/problem_sets/07/data/plants.csv", header = TRUE)
#amel <- data$Amelanchier.canadensis
#Temp <- scale(data[,2])

```


**Exercise 0_a**

Is there a statistically significant relationship between mean temperatures and flowering date? (Note that, when doing a Bayesian analysis, we are not strictly involved in “hypothesis testing” (e.g., we have no null model, no p-values, etc.). However, in a loose mapping between Bayesian approaches and frequentists approaches, data analysts will often use overlap between the credible interval and zero [or other null value] as a measure of “statistical significance”.)

Plot raw, posterior median, 95% percentile of the posterior we don't tend to put prediction interval on plots, we plot posterior mean or median, then interval around that (representing uncertainty about the mean/median line that we're plotting. what do we take the mean of? (for each temp value, we take all of the samples x for alpha[i] + beta[i]*sample, run through all xs, and take median for all of those posteriors. ) 

we don't have to do anything with prediction intervals (these would be applicable to a time-series forecast)

1. raw data + posterior median 
2. histogram of posteriors (alpha, beta, sigma posteriors)
3. chains

use quantile on posteriors to get (e.g. quantile(beta, 0.025, 0.975))
```{r}

data <- read.csv(file = "C:/Users/wmichael/Google Drive/BEE569/problem_sets/07/data/plants.csv", header = TRUE)
data <- data[,c(1,2,3)]


#modeling distributions with normal distribution
#Model for SINGLE species
fileForJAGS_beta <- "Week7_JAGS_Ex0.jags"
cat("
  model {
  for (i in 1:length(y)){
    y[i] ~ dnorm(mu[i], tau)
    mu[i]<- alpha + beta * MAMTemp[i]
  }
  
  alpha ~ dnorm(0,0.001)
  beta ~ dnorm(0,0.001)
  
  tau <- pow(sigma,-2)
  sigma ~ dgamma(0.001,0.001)

    }",fill = TRUE, file= fileForJAGS_beta)

#Data as list for JAGS
Dat <- list(y = data$Amelanchier.canadensis, MAMTemp = data$MAMTemp)

#Function for initial parameter values
InitStage <- function() {list(alpha = rnorm(1,0,1), beta = rnorm(1,0,1), sigma = rgamma(1,2,0.01))} 

#Parameters we want to track
ParsStage <- c("alpha","beta","sigma")

#Variables for MCMC
ni <- 11000  # number of draws from the posterior
nt <- 1    #thinning rate
nb <- 1000  # number to discard for burn-in
nc <- 3  # number of chains


#JAGS model
MCMC_Data_ex0 = jags(inits=InitStage,
         n.chains=nc,
         model.file="Week7_JAGS_Ex0.jags",
         working.directory=getwd(),
         data=Dat,
         parameters.to.save=ParsStage,
         n.thin=nt,
         n.iter=ni,
         n.burnin=nb,
         DIC=T)



#Summarize the model 
MCMCsummary(MCMC_Data_ex0, round = 2)

#Summarize Alpha
MCMCsummary(MCMC_Data_ex0, 
            params = 'alpha')

#Summarize Beta
mc.beta <- MCMCsummary(MCMC_Data_ex0, 
            params = 'beta')



#Check posteriors for convergence
MCMCtrace(MCMC_Data_ex0, params = c('alpha', 'beta', 'sigma'), ISB = FALSE, pdf = FALSE )

#Plot posterior medians, 50%, and 95% credible intervals 
par(mfrow=c(1,1))
MCMCplot(MCMC_Data_ex0, 
         params = 'beta', 
         rank = TRUE,
         xlab = 'ESTIMATE',
         guide_lines = TRUE)

#Plot posterior medians, 50%, and 95% credible intervals 
par(mfrow=c(1,1))
MCMCplot(MCMC_Data_ex0, 
         params = 'alpha', 
         rank = TRUE,
         xlab = 'ESTIMATE',
         guide_lines = TRUE)



alpha.mean<-MCMC_Data_ex0$BUGSoutput$summary[row.names(MCMC_Data_ex0$BUGSoutput$summary)=="alpha",1]; print(alpha.mean)
alpha.median<-MCMC_Data_ex0$BUGSoutput$summary[row.names(MCMC_Data_ex0$BUGSoutput$summary)=="alpha",5]; print(alpha.median)

beta.mean<-MCMC_Data_ex0$BUGSoutput$summary[row.names(MCMC_Data_ex0$BUGSoutput$summary)=="beta",1]; print(beta.mean)
beta.median<-MCMC_Data_ex0$BUGSoutput$summary[row.names(MCMC_Data_ex0$BUGSoutput$summary)=="beta",5]; print(beta.median)

alpha.sd<-MCMC_Data_ex0$BUGSoutput$summary[row.names(MCMC_Data_ex0$BUGSoutput$summary)=="alpha",2]; print(alpha.sd)
beta.sd<-MCMC_Data_ex0$BUGSoutput$summary[row.names(MCMC_Data_ex0$BUGSoutput$summary)=="beta",2]; print(beta.sd)




# #plot the posteriors for alpha and beta
par(mfrow = c(3,1))
hist(MCMC_Data_ex0$BUGSoutput$sims.matrix[,1], breaks = 20, main = NULL, xlab = expression(paste("Posterior Distribution of ", alpha, sep = "")))
hist(MCMC_Data_ex0$BUGSoutput$sims.matrix[,2], breaks = 20, main = NULL, xlab = expression(paste("Posterior Distribution of ", beta, sep = "")))
hist(MCMC_Data_ex0$BUGSoutput$sims.matrix[,3], breaks = 20, main = NULL, xlab = expression(paste("Posterior Distribution of ", sigma, sep = "")))
# 
# #Extract the means (these should be close the the parameter values we initially set (i.e. 2,5)) 
#mean(m$BUGSoutput$sims.matrix[,1])
# mean(m$BUGSoutput$sims.matrix[,2])

# 
# data <- read.csv(file = "C:/Users/wmichael/Google Drive/BEE569/problem_sets/07/data/plants.csv", header = TRUE)
# combined <- gather(data, key = "species", value = "Day", `Amelanchier.canadensis`:`Viola.pedata`)
# p <- qplot(Year, Day,data = combined)
# p + geom_abline(intercept = MCMC_Data_ex0$BUGSoutput$mean$alpha, slope = MCMC_Data_ex0$BUGSoutput$mean$beta, colour = "blue") +
#   geom_abline(intercept = MCMC_Data_ex0$BUGSoutput$summary[9], slope = MCMC_Data_ex0$BUGSoutput$summary[10], colour = "red", linetype = "dashed") +
#   geom_abline(intercept = MCMC_Data_ex0$BUGSoutput$summary[25], slope = MCMC_Data_ex0$BUGSoutput$summary[26], colour = "red", linetype = "dashed")



#plot over original data
data <- read.csv(file = "C:/Users/wmichael/Google Drive/BEE569/problem_sets/07/data/plants.csv", header = TRUE)
data[,1]<- (data[,1] - mean(data[,1]))/sd(data[,1]) #standardizing year
data_long <- melt(data, id.vars = c("Year", "MAMTemp"))

slope_temp <- as.numeric(MCMC_Data_ex0$BUGSoutput$mean[2])
intercept_temp <- as.numeric(MCMC_Data_ex0$BUGSoutput$mean[1])

slope_temp_5 <- as.numeric(MCMC_Data_ex0$BUGSoutput$summary[2,3])
slope_temp_95 <- as.numeric(MCMC_Data_ex0$BUGSoutput$summary[2,7])


intercept_temp_5 <- as.numeric(MCMC_Data_ex0$BUGSoutput$summary[1,3])
intercept_temp_95 <- as.numeric(MCMC_Data_ex0$BUGSoutput$summary[1,7])


ggplot(data_long, aes(x= Year, y = value))+
  geom_point() +
  geom_abline(mapping = aes(slope = slope_temp, intercept = intercept_temp)) +
  geom_abline(mapping = aes(slope = slope_temp_5, intercept = intercept_temp_5), linetype = "dashed") + 
  geom_abline(mapping = aes(slope = slope_temp_5, intercept = intercept_temp_95), linetype = "dashed")



```




**Ex:0b**


```{r}

data <- read.csv(file = "C:/Users/wmichael/Google Drive/BEE569/problem_sets/07/data/plants.csv", header = TRUE)
data[,2]<- (data[,2] - mean(data[,2]))/sd(data[,2]) #standardizing temp


#Model for MULTIPLE species
 fileForJAGS <- "Week7_JAGS_Ex0_temp.jags"
cat("
  model {
for (i in 1:30)
{
    mu[i]<- alpha + beta * MAMTemp[i]
  
for(j in 3:34)
  {
    y[i,j] ~ dnorm(mu[i], tau)
  } #i

}#j
    alpha ~ dnorm(0,0.001)
    beta ~ dnorm(0,0.001)
    
    tau <- pow(sigma,-2)
    sigma ~ dgamma(0.001,0.001)

    }",fill = TRUE, file= fileForJAGS) 
    


#Data as list for JAGS
Dat <- list(y = as.matrix(data), MAMTemp = data[,2])

#Function for initial parameter values
InitStage <- function() {list(alpha = rnorm(1,0,1), beta = rnorm(1,0,1), sigma = rgamma(1,2,0.01))} 

#Parameters we want to track
ParsStage <- c("alpha","beta","sigma")

#Variables for MCMC
ni <- 10000  # number of draws from the posterior
nt <- 1    #thinning rate
nb <- 1000  # number to discard for burn-in
nc <- 3  # number of chains


#JAGS model
mcmc_temp = jags(inits=InitStage,
            n.chains=nc,
            model.file="Week7_JAGS_Ex0_temp.jags",
            working.directory=getwd(),
            data=Dat,
            parameters.to.save=ParsStage,
            n.thin=nt,
            n.iter=ni,
            n.burnin=nb,
            DIC=T)


#Summarize the model 
MCMCsummary(mcmc_temp, round = 2)

#Summarize Alpha
MCMCsummary(mcmc_temp, 
            params = 'alpha')

#Summarize Beta
mc.beta <- MCMCsummary(mcmc_temp, 
            params = 'beta')



#Check posteriors for convergence
MCMCtrace(mcmc_temp, params = c('alpha', 'beta', 'sigma'), ISB = FALSE, pdf = FALSE )

#Plot posterior medians, 50%, and 95% credible intervals 
par(mfrow=c(1,1))
MCMCplot(mcmc_temp, 
         params = 'alpha', 
         rank = TRUE,
         xlab = 'ESTIMATE',
         guide_lines = TRUE)


#Plot posterior medians, 50%, and 95% credible intervals 
par(mfrow=c(1,1))
MCMCplot(mcmc_temp, 
         params = 'beta', 
         rank = TRUE,
         xlab = 'ESTIMATE',
         guide_lines = TRUE)




alpha.mean<-mcmc_temp$BUGSoutput$summary[row.names(mcmc_temp$BUGSoutput$summary)=="alpha",1]; print(alpha.mean)
alpha.median<-mcmc_temp$BUGSoutput$summary[row.names(mcmc_temp$BUGSoutput$summary)=="alpha",5]; print(alpha.median)

beta.mean<-mcmc_temp$BUGSoutput$summary[row.names(mcmc_temp$BUGSoutput$summary)=="beta",1]; print(beta.mean)
beta.median<-mcmc_temp$BUGSoutput$summary[row.names(mcmc_temp$BUGSoutput$summary)=="beta",5]; print(beta.median)

alpha.sd<-mcmc_temp$BUGSoutput$summary[row.names(mcmc_temp$BUGSoutput$summary)=="alpha",2]; print(alpha.sd)
beta.sd<-mcmc_temp$BUGSoutput$summary[row.names(mcmc_temp$BUGSoutput$summary)=="beta",2]; print(beta.sd)




# #plot the posteriors for alpha and beta
par(mfrow = c(3,1))
hist(mcmc_temp$BUGSoutput$sims.matrix[,1], breaks = 20, main = NULL, xlab = expression(paste("Posterior Distribution of ", alpha, sep = "")))
hist(mcmc_temp$BUGSoutput$sims.matrix[,2], breaks = 20, main = NULL, xlab = expression(paste("Posterior Distribution of ", beta, sep = "")))
hist(mcmc_temp$BUGSoutput$sims.matrix[,3], breaks = 20, main = NULL, xlab = expression(paste("Posterior Distribution of ", sigma, sep = "")))


#Plot on the original data with confidence intervals
data <- read.csv(file = "C:/Users/wmichael/Google Drive/BEE569/problem_sets/07/data/plants.csv", header = TRUE)
data[,1]<- (data[,1] - mean(data[,1]))/sd(data[,1]) #standardizing year
combined <- gather(data, key = "species", value = "Day", `Amelanchier.canadensis`:`Viola.pedata`)
p <- qplot(Year, Day, data = combined)
p + geom_abline(intercept = mcmc_temp$BUGSoutput$mean$alpha, slope = mcmc_temp$BUGSoutput$mean$beta, colour = "blue") +
  geom_abline(intercept = mcmc_temp$BUGSoutput$summary[9], slope = mcmc_temp$BUGSoutput$summary[10], colour = "red", linetype = "dashed") +
  geom_abline(intercept = mcmc_temp$BUGSoutput$summary[25], slope = mcmc_temp$BUGSoutput$summary[26], colour = "red", linetype = "dashed")

```

**Ex1 - Year**
Modify the code to use Year rather than Mean Spring Temperature as the covariate. Is there a statistically significant relationship between Year and Flowering Date? Copy and paste code, summary of parameters, and relevant plots.
```{r}

data <- read.csv(file = "C:/Users/wmichael/Google Drive/BEE569/problem_sets/07/data/plants.csv", header = TRUE)
data[,2]<- (data[,2] - mean(data[,2]))/sd(data[,2]) #standardizing temp


 fileForJAGS <- "Week7_JAGS_Ex1_year.jags"
cat("
  model {
    for (i in 1:30)
    {
        mu[i]<- alpha + beta * year[i]
        
    for(j in 3:32)
      {
        y[i,j] ~ dnorm(mu[i], tau)
        
      } #i
    
    }#j
        alpha ~ dnorm(0,0.001)
        beta ~ dnorm(0,0.001)
        
        tau <- pow(sigma,-2)
        sigma ~ dgamma(0.001,0.001)

    }",fill = TRUE, file= fileForJAGS) 
    


#Data as list for JAGS
Dat <- list(y = as.matrix(data), year = data[,1])

#Function for initial parameter values
InitStage <- function() {list(alpha = rnorm(1,0,1), beta = rnorm(1,0,1), sigma = rgamma(1,2,0.01))} 

#Parameters we want to track
ParsStage <- c("alpha","beta","sigma")

#Variables for MCMC
ni <- 10000  # number of draws from the posterior
nt <- 1    #thinning rate
nb <- 1000  # number to discard for burn-in
nc <- 3  # number of chains


#JAGS model
m_year = jags(inits=InitStage,
            n.chains=nc,
            model.file="Week7_JAGS_Ex1_year.jags",
            working.directory=getwd(),
            data=Dat,
            parameters.to.save=ParsStage,
            n.thin=nt,
            n.iter=ni,
            n.burnin=nb,
            DIC=T)


#Summarize the model 
MCMCsummary(m_year, round = 2)

#Summarize Alpha
MCMCsummary(m_year, 
            params = 'alpha')

#Summarize Beta
mc.beta <- MCMCsummary(m_year, 
            params = 'beta')



#Check posteriors for convergence
MCMCtrace(m_year, params = c('alpha', 'beta', 'sigma'), ISB = FALSE, pdf = FALSE )



#Plot posterior medians, 50%, and 95% credible intervals 
par(mfrow=c(1,1))
MCMCplot(m_year, 
         params = 'alpha', 
         rank = TRUE,
         xlab = 'ESTIMATE',
         guide_lines = TRUE)


#Plot posterior medians, 50%, and 95% credible intervals 
par(mfrow=c(1,1))
MCMCplot(m_year, 
         params = 'beta', 
         rank = TRUE,
         xlab = 'ESTIMATE',
         guide_lines = TRUE)




alpha.mean<-m_year$BUGSoutput$summary[row.names(m_year$BUGSoutput$summary)=="alpha",1]; print(alpha.mean)

beta.mean<-m_year$BUGSoutput$summary[row.names(m_year$BUGSoutput$summary)=="beta",1]; print(beta.mean)

alpha.sd<-m_year$BUGSoutput$summary[row.names(m_year$BUGSoutput$summary)=="alpha",2]; print(alpha.sd)
beta.sd<-m_year$BUGSoutput$summary[row.names(m_year$BUGSoutput$summary)=="beta",2]; print(beta.sd)




# #plot the posteriors for alpha and beta
par(mfrow = c(2,1))
hist(m_year$BUGSoutput$sims.matrix[,1], breaks = 20, main = NULL, xlab = expression(paste("Posterior Distribution of ", alpha, sep = "")))
hist(m_year$BUGSoutput$sims.matrix[,2], breaks = 20, main = NULL, xlab = expression(paste("Posterior Distribution of ", beta, sep = "")))


data <- read.csv(file = "C:/Users/wmichael/Google Drive/BEE569/problem_sets/07/data/plants.csv", header = TRUE)
data[,1]<- (data[,1] - mean(data[,1]))/sd(data[,1]) #standardizing year

data_long <- melt(data, id.vars = c("Year", "MAMTemp"))
slope_temp <- as.numeric(m_year$BUGSoutput$mean[2])
intercept_temp <- as.numeric(m_year$BUGSoutput$mean[1])

slope_temp_5 <- as.numeric(m_year$BUGSoutput$summary[2,3])
slope_temp_95 <- as.numeric(m_year$BUGSoutput$summary[2,7])


intercept_temp_5 <- as.numeric(m_year$BUGSoutput$summary[1,3])
intercept_temp_95 <- as.numeric(m_year$BUGSoutput$summary[1,7])


ggplot(data_long, aes(x= Year, y = value))+
  geom_point() +
  geom_abline(mapping = aes(slope = slope_temp, intercept = intercept_temp)) +
  geom_abline(mapping = aes(slope = slope_temp_5, intercept = intercept_temp_5), linetype = "dashed") + 
  geom_abline(mapping = aes(slope = slope_temp_5, intercept = intercept_temp_95), linetype = "dashed")


```
**Ex: 02**

We are now going to collapse the data into the early time period (<1950) and the late time period (≥1950). Year goes from a continuous variable to a discrete (0/1) variable, and the coefficient takes a different interpretation (the difference between flowering times early and late).  Is there a statistically significant shift between the early and late periods? Copy and paste code, summary of parameters, and relevant plots.

Note that this is basically a t-test. Compare your results/interpretation to that obtained using the R function t.test. Interpret in the context of alpha and alpha + beta, interpret posterior of beta, is the posterior very different from zero? T.test gives effect size, but the role of the p-value is played by the location of the posterior, if the posterior is shifted from zero, then you have an effect. 


```{r}

#Run a T.test
data_years <- read.csv(file = "C:/Users/wmichael/Google Drive/BEE569/problem_sets/07/data/plants.csv", header = TRUE)


data_years$Year[data_years$Year < 1950] <- 0
data_years$Year[data_years$Year > 1950] <- 1

year <- data_years$Year
year <- rep(as.numeric(year),times = 32)

y <- unlist(c(data_years[,3:ncol(data_years)])) #
t.test(y~year)
```


```{r}

library(reshape2)

data_years <- read.csv(file = "C:/Users/wmichael/Google Drive/BEE569/problem_sets/07/data/plants.csv", header = TRUE)

data_years$Year[data_years$Year < 1950] <- 0
data_years$Year[data_years$Year > 1950] <- 1
data_years$Year <- factor(data_years$Year)

#Coerce the dataframe into LONG format
data_years.long <- melt(data_years, id.vars=c("Year", "MAMTemp"))

#run the t.test (result should be CI of 5.39 to 9.75)
t.test(data_years.long$value ~ data_years.long$Year)


# ggplot(data_years.long, aes(x=data_years.long$MAMTemp , y= data_years.long$value))+
#   geom_point() +
#   geom_abline(mapping=aes(slope= slope_temp, intercept=  intercept_temp))


 fileForJAGS <- "Week7_JAGS_Ex2_years.jags"
cat("
  model {
for (i in 1:30)
{
    mu[i]<- alpha + beta * year[i]
  
for(j in 3:34)
  {
    y[i,j] ~ dnorm(mu[i], tau)
  }

}
    alpha ~ dnorm(0,0.001)
    beta ~ dnorm(0,0.001)
    
    tau <- pow(sigma,-2)
    sigma ~ dgamma(0.001,0.001)

    }",fill = TRUE, file= fileForJAGS) 
    

#Data as list for JAGS
Dat <- list(y = as.matrix(data_years), year = data_years[,1])

#Function for initial parameter values
InitStage <- function() {list(alpha = rnorm(1,0,1), beta = rnorm(1,0,1), sigma = rgamma(1,2,0.01))} 

#Parameters we want to track
ParsStage <- c("alpha","beta","sigma")

#Variables for MCMC
ni <- 10000  # number of draws from the posterior
nt <- 1    #thinning rate
nb <- 1000  # number to discard for burn-in
nc <- 3  # number of chains


#JAGS model
m_years.factors = jags(inits=InitStage,
            n.chains=nc,
            model.file="Week7_JAGS_Ex2_years.jags",
            working.directory=getwd(),
            data=Dat,
            parameters.to.save=ParsStage,
            n.thin=nt,
            n.iter=ni,
            n.burnin=nb,
            DIC=T)

m_years.factors$BUGSoutput

#Summarize the model 
MCMCsummary(m_years.factors, round = 2)

#Summarize Alpha
MCMCsummary(m_years.factors, 
            params = 'alpha')

#Summarize Beta
mc.beta <- MCMCsummary(m_years.factors, 
            params = 'beta')


#Check posteriors for convergence
MCMCtrace(m_years.factors, params = c('alpha', 'beta', 'sigma'), ISB = FALSE, pdf = FALSE )


#Plot posterior medians, 50%, and 95% credible intervals 
par(mfrow=c(1,1))
MCMCplot(m_years.factors, 
         params = 'alpha', 
         rank = TRUE,
         xlab = 'ESTIMATE',
         guide_lines = TRUE)

#Plot posterior medians, 50%, and 95% credible intervals 
par(mfrow=c(1,1))
MCMCplot(m_years.factors, 
         params = 'beta', 
         rank = TRUE,
         xlab = 'ESTIMATE',
         guide_lines = TRUE)



#Extract means and mediauns
alpha.mean<-m_years.factors$BUGSoutput$summary[row.names(m_years.factors$BUGSoutput$summary)=="alpha",1]; print(alpha.mean)
alpha.median<-m_years.factors$BUGSoutput$summary[row.names(m_years.factors$BUGSoutput$summary)=="alpha",5]; print(alpha.median)

beta.mean<-m_years.factors$BUGSoutput$summary[row.names(m_years.factors$BUGSoutput$summary)=="beta",1]; print(beta.mean)
beta.median<-m_years.factors$BUGSoutput$summary[row.names(m_years.factors$BUGSoutput$summary)=="beta",5]; print(beta.median)

alpha.sd<-m_years.factors$BUGSoutput$summary[row.names(m_years.factors$BUGSoutput$summary)=="alpha",2]; print(alpha.sd)
beta.sd<-m_years.factors$BUGSoutput$summary[row.names(m_years.factors$BUGSoutput$summary)=="beta",2]; print(beta.sd)



# #plot the posteriors for alpha and beta
par(mfrow = c(2,1))
hist(m_years.factors$BUGSoutput$sims.matrix[,1], breaks = 20, main = NULL, xlab = expression(paste("Posterior Distribution of ", alpha, sep = "")))
hist(m_years.factors$BUGSoutput$sims.matrix[,2], breaks = 20, main = NULL, xlab = expression(paste("Posterior Distribution of ", beta, sep = "")))


data <- read.csv(file = "C:/Users/wmichael/Google Drive/BEE569/problem_sets/07/data/plants.csv", header = TRUE)
data[,1]<- (data[,1] - mean(data[,1]))/sd(data[,1]) #standardizing year
data_long <- melt(data, id.vars = c("Year", "MAMTemp"))
slope_temp <- as.numeric(m_years.factors$BUGSoutput$mean[2])
intercept_temp <- as.numeric(m_years.factors$BUGSoutput$mean[1])

slope_temp_5 <- as.numeric(m_years.factors$BUGSoutput$summary[2,3])
slope_temp_95 <- as.numeric(m_years.factors$BUGSoutput$summary[2,7])


intercept_temp_5 <- as.numeric(m_years.factors$BUGSoutput$summary[1,3])
intercept_temp_95 <- as.numeric(m_years.factors$BUGSoutput$summary[1,7])


ggplot(data_long, aes(x= Year, y = value))+
  geom_point() +
  geom_abline(mapping = aes(slope = slope_temp, intercept = intercept_temp)) +
  geom_abline(mapping = aes(slope = slope_temp_5, intercept = intercept_temp_5), linetype = "dashed") + 
  geom_abline(mapping = aes(slope = slope_temp_5, intercept = intercept_temp_95), linetype = "dashed")



```


**Exercise 3**

Tweak the code from Part 1 to include a random effect for species in the α parameter 

```{r}

data <- read.csv(file = "C:/Users/wmichael/Google Drive/BEE569/problem_sets/07/data/plants.csv", header = TRUE)
data[,1]<- (data[,1] - mean(data[,1]))/sd(data[,1]) #standardizing year
data <- melt(data, id.vars=c("Year", "MAMTemp"))
data$variable <- as.numeric(data$variable)
#adapting code from mcarthy box 6.3 for random effects in alpha (intercept)
fileForJAGS <- "Week7_JAGS_year_Ex03.jags"
cat("
  model {
  
#deviation from global alpha for each species
for (i in 1:32){
alpha[i] ~ dnorm(0,tau_among)
}


for (i in 1:960){
 mu[i] <- alpha[species[i]] + beta * year[i]
 y[i] ~dnorm(mu[i], tau_within)
}


beta ~ dnorm(0,0.001)


#Priors for sd and tau among species
tau_among <- pow(sd_among, -2) 
sd_among ~dunif(0,100)


#Priors for sd and tau within species
tau_within <- pow(sd_within, -2)
sd_within ~ dunif(0,100)
}",fill = TRUE, file= fileForJAGS) 
#Data as list for JAGS

Dat <- list(y = data$value, 
            species = data$variable, 
            year = data$Year)

#Function for initial parameter values
InitStage <- function() {list(alpha = rnorm(32,0,1), 
                              beta = rnorm(1,0,1),
                              sd_among = runif(1,0,100), 
                              sd_within = runif(1,0,100))} 

#Parameters we want to track
ParsStage <- c("alpha", "beta", "sd_among", "sd_within")

#Variables for MCMC
ni <- 10000  # number of draws from the posterior
nt <- 1    #thinning rate
nb <- 1000  # number to discard for burn-in
nc <- 3  # number of chains
#JAGS model

MCMC_Year_Ex4 = jags(inits=InitStage,
                    n.chains=nc,
                    model.file="Week7_JAGS_year_Ex03.jags",
                    working.directory=getwd(),
                    data=Dat,
                    parameters.to.save=ParsStage,
                    n.thin=nt,
                    n.iter=ni,
                    n.burnin=nb,
                    DIC=T)




#Summarize the model 
MCMCsummary(MCMC_Year_Ex4, round = 2)

#Summarize Alpha
MCMCsummary(MCMC_Year_Ex4, 
            params = 'alpha')



#Check posteriors for convergence
MCMCtrace(MCMC_Year_Ex4, params = c('alpha', 'sd_among', 'sd_within'), ISB = FALSE, pdf = FALSE )



#Plot posterior medians, 50%, and 95% credible intervals 
par(mfrow=c(1,1))
MCMCplot(MCMC_Year_Ex4, 
         params = 'alpha', 
         rank = TRUE,
         xlab = 'ESTIMATE',
         guide_lines = TRUE)

#Plot posterior medians, 50%, and 95% credible intervals 
par(mfrow=c(1,1))
MCMCplot(MCMC_Year_Ex4, 
         params = 'sd_among', 
         rank = TRUE,
         xlab = 'ESTIMATE',
         guide_lines = TRUE)

#Plot posterior medians, 50%, and 95% credible intervals 
par(mfrow=c(1,1))
MCMCplot(MCMC_Year_Ex4, 
         params = 'sd_within', 
         rank = TRUE,
         xlab = 'ESTIMATE',
         guide_lines = TRUE)









data <- read.csv(file = "C:/Users/wmichael/Google Drive/BEE569/problem_sets/07/data/plants.csv", header = TRUE)
data[,1]<- (data[,1] - mean(data[,1]))/sd(data[,1]) #standardizing year
data_long <- melt(data, id.vars = c("Year", "MAMTemp"))
slope_temp <- as.numeric(m_years.factors$BUGSoutput$mean[2])
intercept_temp <- as.numeric(m_years.factors$BUGSoutput$mean[1])

slope_temp_5 <- as.numeric(m_years.factors$BUGSoutput$summary[2,3])
slope_temp_95 <- as.numeric(m_years.factors$BUGSoutput$summary[2,7])


intercept_temp_5 <- as.numeric(m_years.factors$BUGSoutput$summary[1,3])
intercept_temp_95 <- as.numeric(m_years.factors$BUGSoutput$summary[1,7])


ggplot(data_long, aes(x= Year, y = value))+
  geom_point() +
  geom_abline(mapping = aes(slope = slope_temp, intercept = intercept_temp)) +
  geom_abline(mapping = aes(slope = slope_temp_5, intercept = intercept_temp_5), linetype = "dashed") + 
  geom_abline(mapping = aes(slope = slope_temp_5, intercept = intercept_temp_95), linetype = "dashed")

```

