---
title: "Problem Set 01"
author: "Michael Wethington"
date: "August 28, 2020"
output:
  html_document: default
  pdf_document: default
---

```{r}
#Load Packages

library(ggplot2)
library(ggExtra)
```



**Exercise 1 **
&nbsp;
&nbsp;

Given:
$$ g(X) = \frac{2}{x^3};\ for \ 1 <= X <= \infty $$
$$ and $$
$$ h(Y) = \frac{3}{y^4};\ for\ 1 <= Y <= \infty $$
What is the joint probability distribution of $(X,Y)$?

Show that: $$ P(X<= Y) = \frac{2}{5} $$


Answer: 

We begin by finding the Joint probability of the two pdfs. Since each pdf is independent of one another, the joint pdf of $(X,Y)$ is found by multiplying them. Where, $$ f(x,y) = g(X)\cdot h(Y)$$

$$ f_{X,Y}(x,y)  = \left\{
\begin{array}{ll}
\frac{2}{x^3}\ \cdot \frac{3}{y^4}\ = \ \frac{6}{x^3 y^4} & \quad (1\le X \le \infty, \ 1\le Y \le \infty) \\
0, \ otherwise
\end{array}
\right. $$

We that we have successfully identified the joint pdf, we need to marginalize each respective random variable $(X,Y)$ through the process of integration.

$$ P(X<= Y) = P(1 \le x \le \infty, \ x \le y\le \infty) \\
\int_{1}^{\infty} \int_{x}^{\infty} \ f(x,y) \ dy \ dx \\
\int_{1}^{\infty} \int_{x}^{\infty} \frac{6}{x^3 y^4}\ dy \ dx \\
\int_{x}^{\infty} \frac{6}{x^3 y^4}\ dy  \\  
\frac{6}{x^3 } \int_{x}^{\infty} \frac{1}{y^4}\ dy \\
\frac{6}{x^3 } \int_{x}^{\infty} y^{ - 4} \ dy  \\
\frac{6}{x^3 } \cdot \ \frac{y^ {- 3}}{-3} \ = \frac{-2}{x^3 x^3} \\
\int_{x}^{\infty} \frac{6}{x^3 y^4} = 0 - (- \frac{2}{x^6}) = \ \frac{2}{x^6}\\
\int_{1}^{\infty} \frac{2}{x^6}dx = - \frac{2}{5x^5}\\
\int_{1}^{\infty} \frac{2}{x^6}dx = 0 - (- \frac{2}{5}) \ = \frac{2}{5} $$

**Exercise 2 **
&nbsp;
&nbsp;

In each case, find the (1) joint distribution $ P(X,Y)$, the two marginal distributions $P(X)$ and $P(Y)$ (2 and 3), and (4) determine if X and Y are independent.

Case 1: (X,Y) are uniformly distributed on the square

$$ -6 \le X \le 6 \\ -6 \le Y \le 6 $$

Answer: 

We are using the uniform distribution and find the area of the square to be 144. 

The Join pdf of $(X,Y)$ 

$$ F(X,Y) = \left\{
        \begin{array}{ll}
            \frac{1}{144} & \quad -6 \le x,  y \le 6  \\
            0, \ otherwise
        \end{array}
    \right.$$

Marginal pdf of X:

$$F_{X}(x) =  \int_{-6}^{6} \frac{1}{144}dy  \\
= \left.\frac{1}{144}y\right|_{-6}^6\\$$


$$ F_{X}(x)  = \left\{
        \begin{array}{ll}
            \frac{1}{12} & \quad -6 \le x\le 6  \\
            0, \ otherwise
        \end{array}
    \right.$$


Marginal pdf of Y:


$$
F_{Y}(y) =  \int_{-6}^{6} \frac{1}{144}dx  \\
= \left.\frac{1}{144}x\right|_{-6}^6\\
$$


$$ F_{Y}(y)  = \left\{
        \begin{array}{ll}
            \frac{1}{12} & \quad -6 \le y\le 6  \\
            0, \ otherwise
        \end{array}
    \right.$$

Independence:

Because the product of the marginal pdfs are equal to the joint pdf, we can assume independence of X and Y.




Case 2: (X,Y) are uniformly distributed on the triangle

$$ Y \le X \le 6 \\ -6 \le Y \le X $$

Answer: 

The vertices of the triangle are $(-6,-6), (6,-6),$ and $(6,6)$. The area of the triangle being $\frac{1}{2}\cdot base\cdot height = \frac{1}{2}\cdot 12 \cdot 12 = 72$ \\

Given area of triangle = 72,  

Joint pdf:

$$ F_{X}(x)  = \left\{
        \begin{array}{ll}
            \frac{1}{72} & \quad  Y \le X \le 6, \  -6 \le Y \le X \\
            0, \ otherwise
        \end{array}
    \right.$$
    
Find the double integral:

$$ \int_{-6}^{y} \int_{x}^{6} \ f(x,y) \ dy \ dx \\
 f_{X}(x) = \int_{-6}^{y} \int_{x}^{6} \ f(x,y) \ dy \\ $$


Marginal pdf of X is found by integrating out Y:

$$\begin{align}
F_{X}(x) = \int_{-6}^{x} \ f(x,y) \ dy \\ 
    \int_{-6}^{x} \ \frac{1}{72} \ dy && \text{definite integral } \\ 
    = \frac{1}{72}y|_{-6}^x \ \ && \text{evaluate limits of integration }\\ 
     = \frac{x+6}{72} \\ 
     
\end{align}$$ 

$$ F_{X}(x)  = \left\{
        \begin{array}{ll}
            \frac{x+6}{72} & \quad  -6 \le x \le 6\\
            0, \ otherwise
        \end{array}
    \right.$$


    
Marginal pdf of X:

$$\begin{align}
F_{Y}(y) = \int_{y}^{6} \ f(x,y) \ dx \\ 
    \int_{y}^{6} \ \frac{1}{72} \ dx && \text{definite integral } \\ 
    = \frac{1}{72}y|_{x}^6 \ \ && \text{evaluate limits of integration }\\ 
     = \frac{x+6}{72} \\ 
     
\end{align}$$ 

$$ F_{Y}(y)  = \left\{
        \begin{array}{ll}
            \frac{y-6}{72} & \quad  -6 \le y \le 6\\
            0, \ otherwise
        \end{array}
    \right.$$
    

Independence:
 Since the joint probability of $(X,Y)$ is not equal to the product of the marginal distributions, X and Y are not independent. \\ 



Case 3: (X,Y) are uniformly distributed on the circle

$$ X^2 + Y^2 \le 36 $$
Area of the circle: $36 \pi$ 


The Joint pdf is


$$ F_{X,Y}(x,y)  = \left\{
        \begin{array}{ll}
            \frac{1}{36 \pi} & \quad  x^2 + y^2 \le 36\\
            0, \ otherwise
        \end{array}
    \right.$$
    
    
Marginal pdf of X:

$$\begin{align}
f_{X}(x) = \frac{y}{36 \pi}|_\sqrt{36- x^2} ^\sqrt{-36+x^2} && \text{evaluate limits of integration } \\
 = \frac{1}{18 \pi}\sqrt{36-x^2}
 \end{align}$$



Marginal pdf of Y:

$$ \begin{align}
f_{X}(x) = \frac{x}{36 \pi}|_\sqrt{36- y^2} ^\sqrt{-36+y^2} && \text{evaluate limits of integration } \\
 = \frac{1}{18 \pi}\sqrt{36-y^2}
 \end{align}$$

Independence: 

Because the values of random variables $X$ and $Y$ directly influence done another, and because the product of $X$ and $Y$ does not equal their joint probability $(X,Y)$, they are not independent.  



Write a function for each of the three joint distributions: 

Exercise 1: 
```{r}
library(ggplot2)
library(ggExtra)

##Exercise 1
df1 <- data.frame(x = runif(1000, min = -6, max = 6), y = runif(1000, -6, 6))
p1 <- ggplot(df1, aes(x, y)) + geom_point() + theme_bw()
ggMarginal(p1, type = "histogram")

```

Exercise 2: 
```{r}
# #Exercise 2
# #replicate

runs <- 1000

#set paths for X and Y
X <-rep(NA, runs)
Y<- rep(NA, runs)


#run the loop
for(i in 1:runs){


  #go until we accept the sample
  while(TRUE){

    #sample x and y
    x <- runif(1, -6, 6)
    y <- runif(1, -6, 6)

    #see if we accept the sample
    if(y <= x){

      #accept
      X[i] = x
      Y[i] = y

      #break, go to the next iteration
      break
    }
  }
}

data <- data.frame(X, Y)
p1 <- ggplot(data, aes(X, Y)) + geom_point() + theme_bw()
ggMarginal(p1, type = "histogram")
```


##Exercise 3

```{r}
#Exercise #3
#replicate
set.seed(110)
sims <- 1000

#set paths for X and Y
X <-rep(NA, sims)
Y<- rep(NA, sims)


#run the loop
for(i in 1:sims){


  #go until we accept the sample
  while(TRUE){

    #sample x and y
    x <- runif(1, -6, 6)
    y <- runif(1, -6, 6)

    #see if we accept the sample
    if(x^2 + y^2 <= 36){

      #accept
      X[i] = x
      Y[i] = y

      #break, go to the next iteration
      break
    }
  }
}

data <- data.frame(X, Y)
p1 <- ggplot(data, aes(X, Y)) + geom_point() + theme_bw()
ggMarginal(p1, type = "histogram")

```

