---
title: "Problem Set 01"
author: "Michael Wethington"
date: "August 28, 2020"
output: html_document
---

```{r}

```

**Exercise 1 **
&nbsp;
&nbsp;

Given:
$$ g(X) = \frac{2}{x^3};\ for \ 1 <= X <= \infty $$
$$ and $$
$$ h(Y) = \frac{3}{Y^4};\ for\ 1 <= Y <= \infty $$
What is the joint probability distribution of $(X,Y)$?

Show that: $$ P(X<= Y) = \frac{2}{5} $$


Answer: 

We begin by finding the Joint probability of the two pdfs. Since each pdf is independent of one another, the joint pdf of $(X,Y)$ is found by multiplying them. Where, $$ f(x,y) = g(X)\cdot h(Y) $$

$$\frac{2}{x^3}\ \cdot \frac{3}{Y^4}\ = \ \frac{6}{x^3 y^4}; \ (1\le x \le \infty, \ 1\le y \le \infty) \\ 0,\  otherwise  $$

We that we have successfully identified the joint pdf, we need to marginalize each respective random variable $(X,Y)$ through the process of integration.

$$ P(X<= Y) = P(1 \le x \le \infty, \ x \le y\le \infty) \\ 
\int_{1}^{\infty} \int_{x}^{\infty} \ f(x,y) \ dy \ dx \\
\int_{1}^{\infty} \int_{x}^{\infty} \frac{6}{x^3 y^4}\ dy \ dx \\
\int_{x}^{\infty} \frac{6}{x^3 y^4}\ dy  \\  
\frac{6}{x^3 } \int_{x}^{\infty} \frac{1}{y^4}\ dy \\
\frac{6}{x^3 } \int_{x}^{\infty} y^{ - 4} \ dy \\
\frac{6}{x^3 } \int_{x}^{\infty} y^{ - 4} \ dy \\
\frac{6}{x^3 } \cdot \ \frac{y^ {- 3}}{-3} \ = \frac{-2}{x^3 x^3} \\
\int_{x}^{\infty} \frac{6}{x^3 y^4} = 0 - (- \frac{2}{x^6}) = \ \frac{2}{x^6}\\
\int_{1}^{\infty} \frac{2}{x^6}dx = - \frac{2}{5x^5}\\
\int_{1}^{\infty} \frac{2}{x^6}dx = 0 - (- \frac{2}{5}) \ = \frac{2}{5}
$$

**Exercise 2 **
&nbsp;
&nbsp;

In each case, find the (1) joint distribution $ P(X,Y)$, the two marginal distributions $P(X)$ and $P(Y)$ (2 and 3), and (4) determine if X and Y are independent.

Case 1: (X,Y) are uniformly distributed on the square

$$ -6 \le X \le 6 \\ -6 \le Y \le 6 $$

Answer: 

The Join pdf of X,Y is 

$$f_{X,Y}(x,y) = \frac{1}{12^2}; \ -6 \le x, y \le 6 \\  $$

Marginal pdf of X:

$$f_{X}(x) =  \int_{-6}^{6} \frac{1}{144}dy  \\
\int_{-6}^{6} \frac{1}{144}dy = \frac{1}{12}; \ -6 \le x \le 6\\$$

Marginal pdf of Y:


$$f_{Y}(y) =  \int_{-6}^{6} \frac{1}{144}dx  \\
\int_{-6}^{6} \frac{1}{144}dx = \frac{1}{12}; \ -6 \le y \le 6\\$$

Independence:

Because the product of the marginal pdfs are equal to the joint pdf, we can assume independence of X and Y.

Case 2: (X,Y) are uniformly distributed on the triangle

$$ Y \le X \le 6 \\ -6 \le Y \le X $$

Answer: 

The vertices of the triangle are $(-6,-6), (6,-6),$ and $(6,6)$. The area of the triangle being $\frac{1}{2}\cdot base\cdot height = \frac{1}{2}\cdot 12 \cdot 12 = 72$ \\

Given area of triangle = 72, our joint pdf is: 


$$f(x,y) = \frac{1}{72}; \ \  Y \le X \le 6, \  -6 \le Y \le X $$

$$ \int_{-3}^{y} \int_{x}^{3} \ f(x,y) \ dy \ dx \\
 f_{X}(x) = \int_{-3}^{y} \int_{x}^{3} \ f(x,y) \ dy \\ 
  \int_{x}^{3} \ f(x,y) \ dy \\ 
    \int_{x}^{3} \ \frac{1}{72} \ dy \\
    lim_{y -> x} + (\frac{1}{72}y) = \frac{1}{72}x \\ 
    lim_{y -> 3} + (\frac{1}{72}y) = \frac{1}{24} \\ 
     = \frac{1}{24} - \frac{1}{72}x \\
    \int_{-3}^{y}(\frac{1}{24} - \frac{1}{72}x )\ dx \\
    = \ \int_{-3}^{y}\frac{1}{24}\ dx - \int_{-3}^{y}\frac{1}{72}x \ dx \\
    \ \int_{-3}^{y}\frac{1}{24}\ dx = \frac{1}{24}y - \frac{1}{8} \\
    \int_{-3}^{y}\frac{1}{72}x dx = \frac{1}{72}y - \frac{x^2}{2} \\$$ 
    
    
    
 



Case 3: (X,Y) are uniformly distributed on the circle

$$ X^2 + Y^2 \le 36 $$


Write a function for each of the three joint distributions: 

Exercise 1: 
```{r}
library(ggplot2)
library(ggExtra)

##Exercise 1
df1 <- data.frame(x = runif(1000, min = -6, max = 6), y = runif(1000, -6, 6))
p1 <- ggplot(df1, aes(x, y)) + geom_point() + theme_bw()
ggMarginal(p1, type = "histogram")

```

Exercise 2: 


